Project Manager's Progress Report

### 2017-03-11 pnsac/newsletter/article_EditorsNotes.tex
Adding and updating content for Jan/Feb 2016 newsletter.
Updating all tex files, images.

\title{Editor's Notes} 
\author{Bruce Grant}

@@

This issue of the Chronicle carries the last of four chapters in
Richard Lodge's story of his time at Rolls-Royce. Taken together, they
make an interesting ethnology of an extinct industrial culture. The
whole series is accessible through our website,
{\color{blue}\texttt{\url{http://www.projectnorthstar.ca}}} (note that
the location of the newsletter archive will change when the new
website goes live) starting with the April 2013 issue of the
Chronicle.

@@

We also have another hair-raising flying story from Tim Timmins. A
series of previous Timmins stories can be found through the same
website, starting with the December 2010 issue.

Motorheads among our readers should look into Ted Devey's extensive
writings on the Rolls-Royce Merlin, starting in the December 2006
issue.

All together, the archived issues are a valuable web resource for
readers with an interest in aviation history or the history of Project
North Star.

>>>>

\title{Editor's Notes} 
\author{Roger Button}

@@

@@

This past year started off as a repeat of the difficult year the
Association experienced in 2014. Consequently there was little news
and less enthusiasm for publishing the Chronicle on any regular
basis. However, as a result of changes to the management of the museum
organization there is a renewed sense of hope that we will be able to
move forward in 2016, as expressed in the Notes from the
President. One of those changes is the appointment of Chris Kitzan as
the new director general of CASM. You will find a message from Chris
in this edition. Chris and his boss Alex Benay, the CEO of the museum
corporation, are enthusiastic supporters of our organization. We are also
commencing a new feature highlighting the members of our Association
and starting off with Bruce Gemmill, our Project Manager. In addition,
we are adding a "Miscellany" feature which lists items of interest to
our members. If any of our readers wish to contribute an article, have
suggestions for future articles, have suggestions for the "Miscellany"
feature, or have any comments about the Chronicle please let me know.

%\vspace{12pt}

### 2017-03-11 pnsac/newsletter/article_Feature1.tex
Updating all tex files, images.

\begin{quotation}
	You have to be careful opening up the panels covering the 
	engines, you never know if a bird or some animal might have made
	its home there.
\end{quotation}

That’s one lesson a group of dedicated volunteers has learned in its
10 years of painstaking work restoring a Royal Canadian Air Force
North Star aircraft at the Canada Aviation and Space Museum in Ottawa.

@@

The aircraft had sat neglected outside the museum for
years. “Essentially, the RCAF flew this one into the Rockcliffe Airport
when it was decommissioned. It sat in a old until Robert Holmgren said
someone should do something about it,” said project manager Bruce
Gemmill.

@@

“We approached the museum about restoring the aircraft. At first, they
didn’t know how to take us,” said Gemmill. “The museum had its own
professional conservation staff but it had never worked with a team of
volunteers.”

Eventually the museum agreed to work with the volunteers who formed
the Project North Star Association of Canada with the late Robert
Holmgren as president. “We became the model for other volunteer groups
now working with the museum,” said the association’s current
president, Richard Lodge.

@@

The RCAF assigned the rst North Stars to 412 Squadron where they
transported VIPs and were used in various transfigurations for
reliable, long-range transport services.

During the Korean War the North Star was used by No. 426 (Thunderbird)
Sqdn. to ferry sup- plies across the Pacific Ocean to Japan (Operation
Hawk: The Korean Airlift, July/August). They would y 599 round trips
over the Pacific and deliver seven million pounds of cargo and 13,000
personnel on return trips. All this was achieved without a fatal
crash.

Most of the air force’s North Stars were declared surplus in the
1970s.

@@

Each engine had to be taken o the aircraft, brought into the museum’s
shops where they were taken apart. Each item was tagged, catalogued 
and usually photographed. “We take a lot of photographs. That’s
how we know how they go back together,” said Lodge.

Engine project leader Garry Dupont said, “The first engine took us about
four and a half years to restore. We’re almost finished the third engine
which as taken us about two years. So we are getting faster.” The
propellers were restored separately by Hope Aero Propeller and
Components Inc. in Mississauga, Ont. All restoration work has to be up
to the professional standards set by the museum.

The group plans to restore the aircraft to its original look, not to
have it y again.

@@

The aircraft itself, 17515, has seen countless hours of work put into
re- storing the cockpit with all its switches and navigation
equipment and a small galley for the crew of seven.

@@

“We know this aircraft ew during the Korean War. In fact, we have 
pictures showing it configured with litters to bring the wounded back to
Canada,” said Gimmell. “We don’t want to put in the litters but we
have people working on creating the rope seats that were used.”

The plan is to paint the aircraft to its post-1965 markings, since it
was painted with the Maple Leaf ag on it.

Still the search goes on to nd new members for the association. “The
important thing is to nd new people to whom we can pass on this 
knowledge and skills we are developing,” said Lodge.

>>>>

\begin{quotation}
	\textit{You have to be careful opening up the panels covering the 
	engines, you never know if a bird or some animal might have made
	its home there.}
\end{quotation}

That's one lesson a group of dedicated volunteers has learned in its
10 years of painstaking work restoring a Royal Canadian Air Force
North Star aircraft at the Canada Aviation and Space Museum in Ottawa.

@@

The aircraft had sat neglected outside the museum for
years. "Essentially, the RCAF flew this one into the Rockcliffe Airport
when it was decommissioned. It sat in a field until Robert Holmgren said
someone should do something about it," said project manager Bruce
Gemmill.

@@

"We approached the museum about restoring the aircraft. At first, they
didn't know how to take us," said Gemmill. "The museum had its own
professional conservation staff but it had never worked with a team of
volunteers."

Eventually the museum agreed to work with the volunteers who formed
the Project North Star Association of Canada with the late Robert
Holmgren as president. "We became the model for other volunteer groups
now working with the museum," said the association's current
president, Richard Lodge.

@@

The RCAF assigned the first North Stars to 412 Squadron where they
transported VIPs and were used in various transfigurations for
reliable, long-range transport services.

During the Korean War the North Star was used by No. 426 (Thunderbird)
Sqdn. to ferry supplies across the Pacific Ocean to Japan (Operation
Hawk: The Korean Airlift, July/August). They would fly 599 round trips
over the Pacific and deliver seven million pounds of cargo and 13,000
personnel on return trips. All this was achieved without a fatal
crash.

Most of the air force's North Stars were declared surplus in the
1970s.

@@

Each engine had to be taken o the aircraft, brought into the museum's
shops where they were taken apart. Each item was tagged, catalogued 
and usually photographed. "We take a lot of photographs. That's
how we know how they go back together," said Lodge.

Engine project leader Garry Dupont said, "The first engine took us about
four and a half years to restore. We're almost finished the third engine
which as taken us about two years. So we are getting faster." The
propellers were restored separately by Hope Aero Propeller and
Components Inc. in Mississauga, Ont. All restoration work has to be up
to the professional standards set by the museum.

The group plans to restore the aircraft to its original look, not to
have it fly again.

@@

The aircraft itself, 17515, has seen countless hours of work put into
restoring the cockpit with all its switches and navigation
equipment and a small galley for the crew of seven.

@@

"We know this aircraft flew during the Korean War. In fact, we have 
pictures showing it configured with litters to bring the wounded back to
Canada," said Gemmill. "We don't want to put in the litters but we
have people working on creating the rope seats that were used."

The plan is to paint the aircraft to its post-1965 markings, since it
was painted with the Maple Leaf flag on it.

Still the search goes on to find new members for the association. "The
important thing is to find new people to whom we can pass on this 
knowledge and skills we are developing," said Lodge.

\textit{Reprinted from the March/April 2014 edition of "Legion Magazine," with the kind permission of Jennifer Morse, General Manager of Canvet Publications, and Tom MacGregor.}

### 2017-03-11 pnsac/newsletter/article_MemberInterview.tex
Updating all tex files, images.

\textit{Bruce Gemmill has been involved with Project North Star since 2005 and
	has exceeded 9,000 volunteer hours. Bruce is presently the Project
	Manager, as well as a member of the board of directors.}

>>>>

\begin{quotation}
	\textit{Bruce Gemmill has been involved with Project North Star since 2005 and	has exceeded 9,000 volunteer hours. Bruce is presently the Project Manager, as well as a member of the board of directors.}
\end{quotation}

1.\textit{What is your background in aviation?}

Before joining Project North Star, I served in the Canadian Forces for
38 years as a communications and electronics officer.  Although I
never flew as air crew, I worked on air force communications and radar
systems, air traffic control, and NORAD air defence.  I served on
several air force bases, including North Bay and Trenton, and my last
air force position was senior communications adviser to the Commander
of Canadian NORAD Region in Winnipeg.

2. \textit{How long have you been involved with Project North Star and how and why did you get involved?}

After I retired from the Canadian Forces in 2005, I wanted to do
something that was technically challenging and interesting.  I'd done
some woodworking and motorcycle restoration, and one day I was riding
past the Aviation Museum and decided on a whim I would ask if there
was an opportunity to volunteer on an aircraft restoration project.
Six months later I started on Project North Star.

%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.5]{fwd_cockpit_complete.eps}
%   %caption of the figure 
%   \caption*{\small \em Forward cockpit -- complete.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:fwd_cockpit_complete}
%\end{figure}

3. \textit{What has been the history of your involvement to date?}

Since I had no specific aircraft maintenance training or experience, I
began working on electronic assemblies that had been removed from the
aircraft cockpit during the early days of the project.  With no test
equipment, manuals or spare parts, the work was mainly cleaning and
repainting the cases and front panels.  As I got more familiar with
the equipment, I learned how to spray paint, and did minor repairs to
wiring.  I also worked on electrical components from the Merlin
engine, such as the magnetos and starter motor.

\begin{figure}[htbp]
	\vspace{2em}
	\centering
	%name of the graphic, without the path AND in EPS format:
	\includegraphics[scale=0.9]{resources/PNS2008-10-26_23-31-42.png}
	%caption of the figure 
	\caption*{\small \em Propeller assembly.}
	%label of the figure, which has to correspond to \ref{}:
	\label{fig:propeller}
\end{figure}

The most demanding job was removing several of the propellers so they could be shipped for restoration.  Years of corrosion had sealed the spinners onto the propeller hubs.  Removal required the use of dental picks and a large sledge hammer!

After helping with restoration work on Engine One, I began working on
the cockpit as a major project.  I spent approximately five years
working with other volunteers, beginning with removal of as much
equipment as possible from the cockpit and crew lounge, including the
control column, rudder pedals and center control console, seats and
headliners.  A lot of time was spent cleaning and repairing corrosion
damage before painting the cockpit and installing new insulation.  I
found it very rewarding when we began installing previously restored
equipment to the "new" cockpit, but also very challenging.

\begin{figure}[htbp]
	\vspace{2em}
	\centering
	%name of the graphic, without the path AND in EPS format:
	\includegraphics[scale=1.3]{resources/NorthStarCockpitDec2011.png}
	%caption of the figure 
	\caption*{\small \em Forward cockpit -- complete.}
	%label of the figure, which has to correspond to \ref{}:
	\label{fig:fwd_cockpit_complete}
\end{figure}

 A lot of equipment and fittings had been removed prior to me joining the project, and not all removal work had been documented.  Detective work was required to search out photographs that showed where some equipment was installed.

4. \textit{What has been the highlight of your involvement?}

Completion of the cockpit and crew lounge has certainly been the most
significant achievement for me personally. The project team has
received high praise from numerous visitors who have commented on the
high quality of the restoration work.  Although we consider the work
done, I do still want to complete the restoration by finding a few
missing radios and instruments that were removed by the RCAF.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.75]{resources/NavigatorRack.png}
   %caption of the figure 
   \caption*{\small \em Navigator rack.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:nav-radio-rack}
\end{figure}

5. \textit{What has been the most challenging part of your involvement?}

Finding spare and missing parts for the aircraft has always been a
challenge.  Our ability to search out possible sources is limited, as
are funds.  We have been lucky to get most parts needed for the Rolls
Royce engines, but many items for the airframe are not easy to come
by.  These include everything from very large and expensive items,
such as the de-icing boots for the wings and tail, to door and window
seals.  Rubber and fabric items are particularly hard to find, since
these items deteriorate with age.  Finding the correct paint has
recently become a major challenge because of stricter environmental
laws.  Our goal of achieving an authentic restoration of the North
Star will only become more difficult with the passage of time.

### 2017-03-11 pnsac/newsletter/article_MessageDGCASM.tex
Updating all tex files, images.

\title{Project Manager's Progress Report}
\subtitle{November 2015}
\author{Bruce Gemmill}

@@

The aircraft spent considerable time outside this year, but since we
are short of volunteers, not as much work was done as hoped.  The
aircraft is still outside because we are waiting for space to be freed
up in the storage hangar.  We also expect the North Star to be moved
outside again in early spring, much sooner than normal, to accommodate
other uses of the storage facility.  This extended time outside has
delayed some work, and may result in additional weather damage.

\section{Nr 4 Engine}
\label{sec:engine_4}

The last engine to undergo restoration was disassembled last fall.
Since the engine shop crew have cleaned, painted and re-assembled much
of the engine on the engine assembly stand. The refurbished crankshaft
has been installed in the engine block, and the connecting rods and
pistons cleaned and polished.  Also completed are the more complex
cylinder heads and valve train assemblies, and work has begun to
restore many ancillary pieces, such as the magnetos and other
electrical components, and accessories.

The auxiliary gearbox was removed from Nacelle 4 and has been
completely rebuilt, along with the generator.  These won’t be
installed on the aircraft until shortly before engine 4 is ready to be
installed.

Work has also begun on the engine frame which holds the 12 cylinder
engine and supports the surrounding cowl panels.  Once the engine
frame is ready, the engine will be transferred to it from the engine
stand.  Because of a lack of space and shortage of qualified
volunteers, we don’t expect to have engine 4 completed before 2017.  A
great deal of work is also needed on the cowl panels.  We hope to
start work on these in the New Year.

\section{Fuselage and Main Cabin}
\label{sec:main_cabin}

Work has continued in the main cabin.  After removing all floor and
wall panels, as well as numerous fittings, the floor was cleaned and
all cargo tiedowns thoroughly treated to remove corrosion.  Several
required removal and repair. The starboard windows were repainted and
new Plexiglas installed, then the windows were re-installed on the
aircraft.  The port windows were then removed and repaired, but had to
be installed prior to completion so the aircraft could stay outside
this summer.  These will need to be removed again this winter and the
new glass installed before being returned to the aircraft.  Some
corrosion damage along the exposed ribs and sub floor has been
completed.  The rear bulkhead required considerable work.  This
involved removing a large corroded floor panel and three rear bulkhead
panels.  New panels were made and are being installed.

The forward bulkhead and forward half of the floor have been painted,
but further painting was stopped by the arrival of cold weather in
October.  The painting will have to be completed next summer.

The forward cargo door was cleaned, painted and new seals installed.
The door was put back on the aircraft to secure it while outside this
summer.  This winter we plan on removing the rear cargo door for
similar work.

The tail section and rear belly compartment underwent some cleaning
and repair.  This work will also continue this winter in the storage
hangar.

\section{Cabin Liners and Troop Seats}
\label{sec:liners_troop_seats}

Most original cabin liners have now been repaired or replaced with new
liners.  The liners still need to be painted and many stencils need to
be applied, as well as many new snap fasteners.

The troop seats were built last year and placed in storage.  Some work
has been done to restore the many fitting needed to install the troop
seats, along with litter straps and tiedowns for litters and cargo.

\section{Planned Restoration Work---2016}
\label{sec:plannedwork}

Over the next year, we will continue to work on Engine 4 and the main
cabin and doors.  The aircraft will be moved outside in the late
winter or early spring to allow setup for a planned Star Trek exhibit.
Work will not begin inside the aircraft until the weather warms up, so
further delays are inevitable.

We will be able to make new wall panels and either replace or restore
the wood floor panels.

We hope to have all main cabin painting done this summer, so that the
wall and floor panels can be installed.

One other work item that may be started is to clean and restore the
four engine nacelles.  Pipes and fittings need to be removed, followed
by a thorough cleaning and a fresh coat of paint.  This work may be
delayed until qualified volunteers become available.

%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.5]{fwd_cockpit_complete.eps}
%   %caption of the figure 
%   \caption*{\small \em Forward cockpit -- complete.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:fwd_cockpit_complete}
%\end{figure}

>>>>

\title{Message from the CASM Director General}
%\subtitle{November 2015}
\author{Chris Kitzan}

@@

\begin{quotation}
	\textit{Chris Kitzan was appointed the new director general of CASM last
	November. Previously he was the chief of staff at Library and Archives
	Canada. Before that he worked as a curator and exhibit manager at what
	was then known as the Canadian Museum of Civilization.}
\end{quotation}

My tenure at the Museum is just beginning, but already I have observed
first-hand the contribution volunteers are making, from my meetings
with members, to my perspective as a father watching you
enthusiastically cater to my eight year old son's interests, to the
communications I have received from you with ideas and offers of
support, and finally from guests who have written thanking the Museum
for experiences they have had interacting with you on the floor.  I
have never worked in an organization that had so many individuals
willing to donate anywhere near the amount of time and energy you do.
Given that, is it fair to want more?  I hope so.  With your support, I
would like to tap a little deeper into your energy and discuss ways we
can further work together to innovate and inspire.

I recognize that increased collaboration between passionate parties
will sometimes create moments of turbulence.  I feel it is a small
price to pay if we are all committed to the ultimate goal --
maximizing the potential of this organization.  That is why, on the
North Star front, Richard and CASM staff are currently trying to find
solutions that allow us to meet our collective objectives.  And why I
hope that these initial conversations lead to increased discussions,
and volunteer/staff collaboration on a wider range of fronts.

On the one hand, the Canadair North Star was a reliable, versatile
Canadian aircraft that provided our country with excellent service on
both the military and civilian fronts for decades.  On the other, it
made such a bloody racket that wherever it went it could not be
ignored.  To my mind it's a worthy symbol of the kind of Museum I
would like to see us continue to develop.  Based on its collection,
its staff and its volunteers CASM has already established a reputation
as a world-class Museum.    That standard cannot be sacrificed.  At
the same time, there is room for us to make more noise; to try new
ways to reach Canadians, catch their attention, and further share our
rich collection and stories with them.

### 2017-03-11 pnsac/newsletter/article_PresReport.tex
Updating all tex files, images.

Since my last Notes from the President in December 2013, restoration
work on the North Star has mostly been suspended. The Museum has
needed extra working space so that exhibits could be made ready for
the commemoration this year of the centenary of the start of the First
World War and the 70th anniversary on June 6 of D-Day in the Second
World War. These preparations have resulted in our normal working
areas being unavailable for North Star restoration work.

\begin{figure}[htbp]
  \vspace{2em}
  \centering
  %name of the graphic, without the path AND in EPS format:
  \includegraphics[scale=0.5]{richard_mtg.eps}
  %caption of the figure 
  \caption*{\small \em Richard at the Quarterly Meeting in April.}
  %label of the figure, which has to correspond to \ref{}:
  \label{fig:richard_mtg}
\end{figure}

During the shut down period some limited work has been continuing. Two
volunteers have been working in the cockpit of the plane and two other
volunteers have been doing some work on \#3 engine reassembly.

While the restoration work has been on hold, we have arranged three
monthly Volunteer Lunches at the Museum to enable the people who
normally work on the plane to meet and socialize. Although we would
all like to be working on the plane, we understand why the Museum has
had to ask us to temporarily suspend our activities. The Volunteer
Lunches are a way of keeping the team together and up-to-date on what
is happening.

At the last meeting of the PNSAC Board of Directors we agreed to
reschedule the Annual General Meeting from June each year to
September. It has always been difficult to find a good weekend in June
to hold the AGM and it is felt that a middle of September date would
be much better. Our corporate secretary, Roger Button, is aware of
this proposed change and has confirmed that the existing Board of
Directors can extend its term until the new AGM date.

Although there has been little restoration work done since December,
we are working with the Museum to get the project back on track again
in early June. We are also working to enable us to recruit new
volunteers to replace the several people who for various reasons have
had to give up active work.

One of the recent discussion areas has been in connection with
long-term funding of the restoration project. Everybody is aware that
funding from the Federal Government has been substantially reduced. We
have accepted that major  expenditures will have to be met from funds
raised by donations from friends of the Museum and members of our
Association. This is a significant challenge since everybody receives
many appeals throughout the year for donations from many different
deserving organizations.

During the next few months we hope to develop a revised restoration
work plan which will enable us to estimate the funds needed to
complete the project. We will then be able to target appeals for funds
to carry out specific parts of the restoration in the same way as we
did for our successful appeal for funds to make troop seats for the
interior of the plane.

>>>>

As I write these words I am watching the falling snow and preparing for a nasty weather mix this evening and at the same time thinking of the spring and summer and our outdoor activities with the North Star aircraft. 2016 has every chance of being an interesting year for the North Star volunteers. Although full details are not available yet, we are expecting to be able to show off the plane to the general public and to other groups several times this year.

We are continuing to work closely with the Museum conservation staff to develop plans for extended hours working which basically means running a restoration shift on one or more Saturdays each month. This should enable some members to become Restoration Volunteers even though they are not available for work Mondays to Fridays. We had hoped to start the shift before Easter but it looks now as if the start date will be in the fall. There are still some details of supervision, security and safety to be worked out. Running a Saturday shift is a new endeavour for the Museum and it is important for everybody that we take the necessary time to get the operation properly setup so that we have a successful start and a good model for any further out of hours working arrangements for PNSAC or any other volunteer organization working on a museum corporation artifact.

We were very pleased to welcome Chris Kitzan, the new Director General of the Canada Aviation and Space Museum to our first Members' Meeting of 2016 which was held in the Bush Theatre on February 27th. Chris acknowledged the importance of the partnership between the museum and Project North Star. Following the meeting, I saw Chris having a lively and upbeat conversation with our members, over coffee and snacks.

Prior to the Members' Meeting, several of us ran into Mike Charters, who is a volunteer maintenance mechanic at the Warplane Heritage Museum in Hamilton. Mike was one of the team who flew to England in the summer of 2014 for the tour of the UK by the Canadian Lancaster together with the British Lancaster. Mike generously agreed to speak to the members for a few minutes and had some very interesting stories about the trip.

One of the best activities for our members is to become an active volunteer. Our volunteers fall into two categories, Restoration Volunteers and Association Volunteers. The former are those who work on the North Star restoration and have to be limited in number because it is not possible to have too many people working at any one time; this is mainly for supervision and safety reasons. Association Volunteers are those who volunteer for other activities in our Association. There are many different things which need to be done in a volunteer organization and we encourage our members to contact us to volunteer for such things as meeting the public on aircraft open days, helping to organize events and assisting with sales of our merchandise. Volunteering has many rewards, from comradeship with other volunteers, enjoying involvement with an interesting aviation artifact and personal satisfaction derived from doing something worthwhile. If you are interested in exploring volunteering opportunities, do contact any one of our directors whose contact information is shown at the end of this newsletter.

### 2017-03-11 pnsac/newsletter/article_ProgReport.tex
Updating all tex files, images.

\title{Project Manager's Progress Report}
\subtitle{April 2014}
\author{Bruce Gemmill}

@@

Due to museum preparations for a display to mark the beginning of
the First World War and other events in 2014, the restoration
space used by Project North Star was taken over by other museum
projects, and most work on the North Star was suspended.
However, some limited work did continue.

\section{Nr 3 Engine}
\label{sec:engines_3}

Work on Engine 3 was halted for several weeks in early 2014, but
has now resumed on a reduced work schedule.  Most accessories are
now installed, and soon the cowl panels will be fitted.  It is
expected that Engine 3 will be installed late this spring or
early summer, so that work can begin on removing and restoring
Engine 4.

This downtime also offered the opportunity for the engine crew to
clean and re-organize the engine shop.

\section{Cockpit}
\label{cockpit}

While restoration work in the shop could not be continued, we
used the opportunity to install the crew seats and other restored
items to the cockpit.  The pilot seat was installed (first
officer seat was installed last year) then the railings and
safety board were installed behind the seats.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{fwd_cockpit_complete.eps}
   %caption of the figure 
   \caption*{\small \em Forward cockpit -- complete.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:fwd_cockpit_complete}
\end{figure}

Next we installed
the radio officer and navigator tables, and the seats for these
positions.  The radio officer seat proved to be a particular
challenge, as the floor in this area is no longer flat, and some
modifications to the seat rails were necessary to allow the seat
to slide.  Finally the flight engineer's jump seat was attached
to the railing.  With five crew seats installed, the cockpit is
very cramped.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{radio_position.eps}
   %caption of the figure 
   \caption*{\small \em Radio Officer's position showing the morse
     code key.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:radio_position}
\end{figure}

The cockpit installations were finished off by installing the
curtains, crew oxygen hoses, and a Morse code key at the radio
position.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{cockpit_curtain.eps}
   %caption of the figure 
   \caption*{\small \em Finishing touches -- the cockpit curtain.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:cockpit_curtain}
\end{figure}

\section{Crew Lounge, Galley,  and Forward Washroom}
\label{crewlounge}

The floor under this section was repaired by removing a section
of badly corroded I-beam and fabricating a new one from several
pieces of aluminum which were riveted in place to support the
floor.  Once this repair was completed, new cushion flooring was
cut and glued to the crew lounge and the washroom floor, then
various pieces of floor trim were installed.  To install the crew
seat, it was necessary to remove the lower support section, since
the fully assembled seat could not easily fit through the narrow
doorway.  Once the seat was installed, we added the crew table
and bunk bed, along with the secure storage box.  We also
installed the curtain rod and new curtains to close off the bunk
bed. The installation was finished off by re-installing the door
between the lounge and the main cabin, so that this fully
restored area can be protected from dust and debris while the
main cabin undergoes restoration over the next two years.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{new_galley_floor.eps}
   %caption of the figure 
   \caption*{\small \em New galley and floor.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:new_galley_floor}
\end{figure}

The completion of the forward area of the aircraft marks a
significant milestone in the restoration of the North Star.

\section{Planned Restoration Work 2014}
\label{sec:plannedwork}

Some limited work is being done to remove fittings and wall
panels from the main cabin, in preparation for some major
restoration work this summer.  Fabric work on the troop seats is
nearing completion, although the seats will not be installed
until all work on the main cabin is complete.  We hope to resume
a full work schedule by June, with the aircraft outside once
more.

>>>>

\title{Project Manager's Progress Report}
\subtitle{November 2015}
\author{Bruce Gemmill}

@@

The aircraft spent considerable time outside this year, but since we
are short of volunteers, not as much work was done as hoped.  The
aircraft is still outside because we are waiting for space to be freed
up in the storage hangar.  We also expect the North Star to be moved
outside again in early spring, much sooner than normal, to accommodate
other uses of the storage facility.  This extended time outside has
delayed some work, and may result in additional weather damage.

\section{Nr 4 Engine}
\label{sec:engine_4}

The last engine to undergo restoration was disassembled last fall.
Since the engine shop crew have cleaned, painted and re-assembled much
of the engine on the engine assembly stand. The refurbished crankshaft
has been installed in the engine block, and the connecting rods and
pistons cleaned and polished.  Also completed are the more complex
cylinder heads and valve train assemblies, and work has begun to
restore many ancillary pieces, such as the magnetos and other
electrical components, and accessories.

The auxiliary gearbox was removed from Nacelle 4 and has been
completely rebuilt, along with the generator.  These won't be
installed on the aircraft until shortly before engine 4 is ready to be
installed.

Work has also begun on the engine frame which holds the 12 cylinder
engine and supports the surrounding cowl panels.  Once the engine
frame is ready, the engine will be transferred to it from the engine
stand.  Because of a lack of space and shortage of qualified
volunteers, we don't expect to have engine 4 completed before 2017.  A
great deal of work is also needed on the cowl panels.  We hope to
start work on these in the New Year.

\section{Fuselage and Main Cabin}
\label{sec:main_cabin}

Work has continued in the main cabin.  After removing all floor and
wall panels, as well as numerous fittings, the floor was cleaned and
all cargo tiedowns thoroughly treated to remove corrosion.  Several
required removal and repair. The starboard windows were repainted and
new Plexiglas installed, then the windows were re-installed on the
aircraft.  The port windows were then removed and repaired, but had to
be installed prior to completion so the aircraft could stay outside
this summer.  These will need to be removed again this winter and the
new glass installed before being returned to the aircraft.  Some
corrosion damage along the exposed ribs and sub floor has been
completed.  The rear bulkhead required considerable work.  This
involved removing a large corroded floor panel and three rear bulkhead
panels.  New panels were made and are being installed.

The forward bulkhead and forward half of the floor have been painted,
but further painting was stopped by the arrival of cold weather in
October.  The painting will have to be completed next summer.

The forward cargo door was cleaned, painted and new seals installed.
The door was put back on the aircraft to secure it while outside this
summer.  This winter we plan on removing the rear cargo door for
similar work.

The tail section and rear belly compartment underwent some cleaning
and repair.  This work will also continue this winter in the storage
hangar.

\section{Cabin Liners and Troop Seats}
\label{sec:liners_troop_seats}

Most original cabin liners have now been repaired or replaced with new
liners.  The liners still need to be painted and many stencils need to
be applied, as well as many new snap fasteners.

The troop seats were built last year and placed in storage.  Some work
has been done to restore the many fitting needed to install the troop
seats, along with litter straps and tiedowns for litters and cargo.

\section{Planned Restoration Work---2016}
\label{sec:plannedwork}

Over the next year, we will continue to work on Engine 4 and the main
cabin and doors.  The aircraft will be moved outside in the late
winter or early spring to allow setup for a planned Star Trek exhibit.
Work will not begin inside the aircraft until the weather warms up, so
further delays are inevitable.

We will be able to make new wall panels and either replace or restore
the wood floor panels.

We hope to have all main cabin painting done this summer, so that the
wall and floor panels can be installed.

One other work item that may be started is to clean and restore the
four engine nacelles.  Pipes and fittings need to be removed, followed
by a thorough cleaning and a fresh coat of paint.  This work may be
delayed until qualified volunteers become available.

\section{March 2016 Update}
\label{sec:update}

Since our report at the November AGM, the aircraft was moved inside,
so that it could properly sealed against the weather, as it will spend
much of 2016 outside. The aircraft was moved out again in March.
During this time, very little work was done on the aircraft, except
for replacement of the Plexiglas in the port windows, and the
installation of the ARC-552 UHF radio, kindly transferred to the
museum by the Department of National Defence.

\subsection{Nr 4 Engine}
\label{sec:update-engine}

The engine frame has been refurbished, and the engine was transferred
to the frame, so that ancillary equipment can be installed.  Work has
also begun on the supercharger, as well as such items as the starter
motor and header tank.  Some work is also being done on the cowl panels.

%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.5]{fwd_cockpit_complete.eps}
%   %caption of the figure 
%   \caption*{\small \em Forward cockpit -- complete.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:fwd_cockpit_complete}
%\end{figure}
